% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
\usepackage{EMNLP2023}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage{float}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out.
% However, it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

% This is also not strictly necessary, and may be commented out.
% However, it will improve the aesthetics of text in
% the typewriter font.
\usepackage{inconsolata}
% includegraphics
\usepackage{graphicx}
%listings
\usepackage{listings}
\usepackage{dirtytalk}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{array}
\usepackage{booktabs}
%review macros
\usepackage{xcolor}
\newcommand{\review}[1]{{\color{black}#1}}

%pandas tables
\usepackage{{booktabs}}
% Commands
\newcommand{\todo}[1]{{\color{red}\colorbox{yellow}{\textbf{TODO: }}#1}}
\newcommand{\averitec}{AVerImaTeC}
\newcommand{\evr}{Ev\textsuperscript{2}R}
\newcommand{\supp}{Supported}
\newcommand{\reff}{Refuted}
\newcommand{\nei}{Not enough evidence}
\newcommand{\conf}{Conflicting evidence/Cherrypicking}
\makeatletter
\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
\makeatother

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{AIC CTU@AVerImaTeC: dual-retriever RAG for image-text fact checking}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{Herbert Ullrich \\
AI Center @ CTU FEE\\
Charles Square 13\\
Prague, Czech Republic\\
\texttt{ullriher@fel.cvut.cz} \\\And
Jan Drchal \\
AI Center @ CTU FEE\\
Charles Square 13\\
Prague, Czech Republic\\
\texttt{drchajan@fel.cvut.cz} \\}

\begin{document}
%{\makeatletter\acl@finalcopytrue
  \maketitle
%}
\begin{abstract}
In this paper, we present our 3rd place system in the AVerImaTeC shared task, which combines our last year's retrieval-augmented generation (RAG) pipeline with a reverse image search (RIS) module.
Despite its simplicity, our system delivers competitive performance with a single multimodal LLM call per fact-check at just \$0.013 on average using GPT5.1 via OpenAI batch API.
Our system is also easy to reproduce and tweak, consisting of only three decoupled modules -- a textual retrieval module based on similarity search, an image retrieval module based on API-accessed RIS, and a generation module using GPT5.1 -- which is why we suggest it as an accesible starting point for further experimentation.
We publish its code\footnote{\url{https://github.com/heruberuto/AVerImaTec\_Shared\_Task}} and prompts, as well as our vector stores and insights into the scheme's running costs and directions for further improvement.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% inputs
\input{src/introduction}
\input{src/system_description}
%\input{src/classification}
\input{src/results}
%\input{src/software.tex}
\input{src/conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Limitations}
Our pipeline is not meant to be relied upon nor to replace a human fact-checker, but rather to assist an informed user. It gives sources for both the textual and image-text evidence and proposes labels for further questioning. Further hallucination could be spotted within the generated justification, however, the system is not meant to be used as an oracle. The current prompting and text-retrieval model assume English input, and the MLLM-backbone used for the shared task (although interchangeable) is GPT5.1, which is a black box model with limited reproducibility and considerable carbon costs. 
Refuted class is massively overrepresented in the \averitec{} dataset (95\% of train-claims and 78\% of text-claims), making the accuracy-based \averitec{} score computed over \averitec{} test set a problematic metric for systems used in the wild.

\section*{Ethics statement}
Our pipeline is an extension of our already existing last year submission all original authors agreed with.
The system was build specifically for the~\averitec~shared task and reflects the biases of its annotators, for more information on this, we suggest the original \averitec{} paper~\cite{cao2025averimatecdatasetautomaticverification}.
\section*{Acknowledgements}
We would like to thank Tomáš Mlynář for providing insights into multi-modal retrieval systems.

This article was created with the state support of the Ministry of Industry and Trade of the Czech Republic, project no. Z220312000000, within the National Recovery Plan Programme.
The access to the computational infrastructure of the OP VVV funded project CZ.02.1.01/0.0/0.0/16\_019/0000765 ``Research Center for Informatics'' is also gratefully acknowledged.



% Entries for the entire Anthology, followed by custom entries
\bibliography{anthology,custom}
\bibliographystyle{acl_natbib}

\appendix

\include{src/appendix_a_llms}
%\include{src/appendix_b_think}
%\include{src/appendix_b_opensource}
%\include{src/appendix_c_errors}

\end{document}

